---
title: "Define binding sites example code"
author: "Melina Klostermann"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-overflow: scroll
    code-summary: "Show code"
    code-tools: true
    code-line-numbers: true
    
    toc: true
    toc-depth: 3
    toc-location: left
    toc-expand: false
    number-sections: true
    
    theme: sandstone
    fontsize: 11pt
    linestretch: 1.5
    fig-format: svg
        
    cap-location: margin
    crossref:
      fig-title: Fig
    
    embed-resources: true
    link-external-newwindow: true
    smooth-scroll: true
    
    execute:
      echo: true
      warning: false
      cache: false

---
  
```{r setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=FALSE, cache.lazy = FALSE) #, fig.pos = "!H", out.extra = ""
```

```{r libraries, include=FALSE}
# ----------------------
# Load libraries
# ----------------------
library(knitr)
library(tidyverse)
library(purrr)

library(rtracklayer)
library(BindingSiteFinder)
library(GenomicRanges)
library(GenomeInfoDb)
library(txdbmaker)
library(Biostrings)
library(BSgenome.Hsapiens.UCSC.hg38)

# set an output folder
out <- "/Users/jb/data/HNPRNPH/additional analysis/bsfinderout/"

set.seed(5)

```

# What is done here?

This is an exemplary code to define binding sites from racoon_clip outputs. Here the replicates of the for the U2AF2 iCLIP3 dataset with 250ug RNA are used. 

# Input files

```{r}
# ----------------------
# Get input from racoon_clip
# ----------------------
# pureclip files
pureclip_file <- "/Users/jb/data/HNPRNPH/Tretow_et_al_2023/data/iCLIP/PureCLIP_HNRNPH_crosslink_sites.bed"


# bw files
bw_dir <- "/Users/jb/data/HNPRNPH/Tretow_et_al_2023/data/iCLIP"
bw.plus <- list.files(bw_dir, pattern = "*\\plus.bw$", full.names = TRUE, recursive = TRUE)
bw.minus <- list.files(bw_dir, pattern = "*\\minus.bw$", full.names = TRUE, recursive = TRUE)
```

```{r eval = FALSE}
# ----------------------
# Prepare annotation
# ----------------------
# Note: run only once to prepare annotation files for BSF and save them as rds files

# annotation file (download for example from GENCODE)
annoFile <- "/Users/jb/data/HNPRNPH/Tretow_et_al_2023/data/gencode.v39.annotation.gtf.gz"
out_gns <- paste0(out, "gns.rds")
out_regions <- paste0(out, "regions.rds")

# prepare annotation for BSF
###########################
# Make annotation database from gff3 file
annoDb = txdbmaker::makeTxDbFromGFF(file = annoFile, format = "gtf")
annoInfo = rtracklayer::import(annoFile, format = "gtf")


# Get genes as GRanges
gns = genes(annoDb)
idx = match(gns$gene_id, annoInfo$gene_id)
meta = cbind(elementMetadata(gns),
             elementMetadata(annoInfo)[idx,])
meta = meta[,!duplicated(colnames(meta))]
elementMetadata(gns) = meta

saveRDS(gns, out_gns)


# Get regions as Granges
cdseq = cds(annoDb) 
intrns = unlist(intronsByTranscript(annoDb)) 
utrs3 = unlist(threeUTRsByTranscript(annoDb)) 
utrs5 = unlist(fiveUTRsByTranscript(annoDb)) 
regions = GRangesList(CDS = cdseq, INTRON = intrns, UTR3 = utrs3, UTR5 = utrs5)

saveRDS(regions, out_regions)
```


```{r}
# read prepared annotation
gns <- readRDS(paste0(out, "gns.rds"))
regions <- readRDS(paste0(out, "regions.rds"))
```


## Pureclip peaks

```{r}
#-----------------
# Peaks from pureclip
#-----------------

# read pureclip peaks
peaks  = rtracklayer::import(con = pureclip_file, 
                             format = "BED", 
                             extraCols=c("additionalScores" = "character"))

# clean pureclip peaks columns
peaks$additionalScores = NULL
peaks$name = NULL

# optionally keep only standard chromosomes and drop chrM
peaks = keepStandardChromosomes(peaks, pruning.mode = "coarse") %>% 
                 dropSeqlevels(., "chrM", pruning.mode = "coarse" )

# check number of peaks
NROW(peaks)
```


## Metadata for BindingSiteFinder

```{r}
# --------------------
# make metadata for BindingSiteFinder
# --------------------

meta = data.frame(
  id = c(1), # give each sample a unique id
  condition = c("250ug"), # add the condition for each sample (used for differential analysis), but one BSF object per condition needs to be used
  clPlus = bw.plus, # crosslinks plus strand bigwig files
  clMinus = bw.minus) # crosslinks minus strand bigwig files


meta
```

# Make binding sites

## Use automatic binding site generation

```{r}
# --------------------
# run BindingSiteFinder in automatic mode
# --------------------

# make BSF objects
bds_object = BSFDataSetFromBigWig(ranges = peaks, 
                           meta = meta, 
                           silent =T)

# make inital BS allowing BSF to determine automatic parameters
bds_automatic = BSFind(bds_object, 
                 anno.genes = gns, 
                 anno.transcriptRegionList = regions, 
                 cutoff.geneWiseFilter = 0)


# save automatic binding sites
saveRDS(bds_automatic, paste0(out, "bds_automatic.rds"))
exportToBED(bds_automatic, con = paste0(out, "binding_sites_automatic.bed"))
```

## Check automatic binding sites (and decide what parameters should be changed for final binding site definition)


```{r fig.width=5, fig.height=5}
# --------------------
# display paramteres choosen by automatic BSF
# --------------------

bds_automatic <- readRDS(paste0(out, "bds_automatic.rds"))

processingStepsFlowChart(bds_automatic)
ggsave(filename = paste0(out, "processing_steps_automatic_bs.pdf"))
```


# Control (and optionally change) the binding site width


```{r fig.width=5, fig.height=5}

estimateBsWidthPlot(bds_automatic)
ggsave(filename = paste0(out, "estimate_binding_site_width.pdf"))
```

```{r}
# --------------------
# Compare different binding site widths
# --------------------

# compute binding sites with different widths
bds1 <- makeBindingSites(object = bds_object, bsSize = 5)
bds2 <- makeBindingSites(object = bds_object, bsSize = 3)
bds3 <- makeBindingSites(object = bds_object, bsSize = 7)
bds4 <- makeBindingSites(object = bds_object, bsSize = 9)

# summarize in list
l = list(`automatic - bsSize = 5` = bds1, `close by - bsSize = 3` = bds2, 
         `close by - bsSize = 7` = bds3, `close by - bsSize = 9` = bds4)
# plot comparison
rangeCoveragePlot(l, width = 20, show.samples = F, subset.chromosome = "chr1")
ggsave(filename = paste0(out, "compare_binding_site_widths.pdf"))

# save bs for igv
exportToBED(bds2, con = paste0(out, "binding_sites_bs3.bed"))
exportToBED(bds3, con = paste0(out, "binding_sites_bs7.bed"))
exportToBED(bds4, con = paste0(out, "binding_sites_bs9.bed"))

```


# Session Info

```{r}
sessionInfo()
```
