---
title: "RTstop-profiling QC"
author: "Mario Keller"
date: "`r format(Sys.time(), '%B %e, %Y')`"
geometry: "left=1cm,right=1cm,top=1.5cm,bottom=1.5cm"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: true
    toc_depth: 3
    number_sections: true
params:
  study: "TCGA-LIHC"
  workers: 3

header-includes:
- \makeatletter\renewcommand*{\fps@figure}{h}\makeatother
- \usepackage{placeins}
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyr)
library(egg)
library(ggplot2)
library(corrplot)
library(ComplexHeatmap)
library(GGally)
library(ggrastr)
library(gridExtra)
library(ggpointdensity)
library(viridis)
library(ggbeeswarm)
library(ggseqlogo)
library(Biostrings)
library(stringr)
library(dplyr)
library(matrixStats)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.pos = 'h', fig.align = "center", fig.width = 6, fig.height=6)
```

```{r ggplot.themes, echo=FALSE, message=FALSE, warning=FALSE}
ggplot.theme <- theme_bw() +
      theme(
        axis.text=element_text(size=7,face="bold", color = "black"),
        axis.title.x=element_text(size=8,face="bold"),
        axis.title.y=element_text(size=8,face="bold", vjust = 0),
        plot.title = element_text(size=8,face="bold"),
        legend.text = element_text(size=8,face="bold"),
        legend.title = element_text(size=8,face="bold"),
        legend.position = "bottom",
        legend.direction="horizontal")

ggplot.theme.no.legend <- theme_bw() +
      theme(
        axis.text=element_text(size=7,face="bold", color = "black"),
        axis.title.x=element_text(size=8,face="bold"),
        axis.title.y=element_text(size=8,face="bold", vjust = 0),
        plot.title = element_text(size=8,face="bold"),
        legend.text = element_text(size=8,face="bold"),
        legend.title = element_text(size=8,face="bold"),
        legend.position = "none",
        legend.direction="vertical")
```

```{r define_cols}
experiment_cols = c("GTP_KCL_rep1" = "#1b8904", "GTP_KCL_rep2" = "#1b8904", "GTP_KCL_rep3" = "#1b8904",
                    "GTP_NaCL_rep1" = "#0a1fb3", "GTP_NaCL_rep2" = "#0a1fb3", "GTP_NaCL_rep3" = "#0a1fb3",
                    "7dGTP_KCL_rep1" = "#91da81", "7dGTP_KCL_rep2" = "#91da81", "7dGTP_KCL_rep3" = "#91da81",
                    "7dGTP_NaCL_rep1" = "#7d8aea", "7dGTP_NaCL_rep2" = "#7d8aea", "7dGTP_NaCL_rep3" = "#7d8aea")
```

\pagebreak

```{r load.data, echo=F, warning=F, message=F}
RTstop.profiling.l <- readRDS("../rds/RTstop.profiling.l.rds")
construct_info.df <- readRDS("../rds/construct.information.df.rds")
# RTstop.profiling.l <- lapply(RTstop.profiling.l, function(l){l[1:100]})
```

# Preprocessing 

Read counts (RCs) at single-nucleotide resolution were generated by importing the BAM-Files and taking the nucleotide upstream 
of the read starts. If a read started at position 1 the RC was assigned to position 1 instead of position 0.

# Library size and normalization

A first look on the library sizes (sum of RCs in a library) revealed for the GTP KCl experiments slightly
more RCs than for the other experiments. 

```{r lib_sizes, fig.width=8, fig.height=3}
RCs_per_construct.df <- lapply(RTstop.profiling.l, sum) %>% do.call("cbind", .) %>% as.data.frame

lib_sizes.df <- RCs_per_construct.df %>% colSums %>% as.data.frame %>% add_rownames %>% dplyr::rename(., lib_size = .) %>% mutate(rowname = factor(rowname, levels = rowname))

lib_sizes.df %>%
  ggplot(., aes(x=rowname, y=lib_size, fill=rowname)) + geom_col(col="black") + scale_fill_manual(values=experiment_cols) + ggplot.theme.no.legend + theme(axis.text.x = element_text(angle=45, vjust=.6)) +
  scale_y_continuous(labels = scales::comma, breaks=seq(0,15000000, by=2500000)) +
  labs(x="Experiment", y="Sum of RCs")
```
\FloatBarrier

Because of these differences in library size, normalization was required. For this purpose, the single-nucleotide RCs were divided
by the library size with a subsequent multiplication by 10^6. The resulting values are called read counts per million (RCPM). 

```{r normalize_RCPM}
RCPMs.l <- lapply(names(RTstop.profiling.l), function(exp){
  RTstop.profiling.l[[exp]] / lib_sizes.df$lib_size[lib_sizes.df$rowname == exp] * 10^6
})
names(RCPMs.l) <- names(RTstop.profiling.l)


RCPMs_per_construct.df <- lapply(RCPMs.l, sum) %>% do.call("cbind", .) %>% as.data.frame

log2_RCPMs_per_construct.df <- (RCPMs_per_construct.df + 1) %>% log2
```

In the following plot the RCPM distribution for the constructs across the 12 experiments is shown.

```{r RCPM_distributions, fig.width=8, fig.height=3}
log2_RCPMs_per_construct.df %>% pivot_longer(cols = everything()) %>% mutate(name = factor(name, levels = names(experiment_cols))) %>% 
  ggplot(., aes(x=name, y=value, fill=name)) + geom_violin(alpha=.5) + geom_boxplot(outlier.size = -1, width=.25, alpha=.75) +
  scale_fill_manual(values=experiment_cols) + 
  labs(x="Experiment", y="log2(RCPM + 1)") + ggplot.theme.no.legend + theme(axis.text.x = element_text(angle=45, vjust=.6))
```

\FloatBarrier
\pagebreak

# Metaprofiles + Heatmap

For a first overview metaprofiles were combined with heatmaps to visualize the distribution RCPM values along the constructs for
all 12 experiments. Rows were ordered by the position with the maximum RCPM value in the range from position 19 to 164.

```{r metaprofile_heatmaps, fig.width=12, fig.height=3}

hm_col <- circlize::colorRamp2(c(0, 5, 10, 20), c(viridis(4)[1], viridis(4)[2], viridis(4)[3], viridis(4)[4]))

heatmaps.l <- lapply(RCPMs.l, function(l){
  m <- l %>% as.matrix
  rownames(m) <- l %>% names
  
  row_order <- m[,19:200] %>% apply(1, which.max) %>% unlist  %>% order %>% rev
  m <- m[row_order,]

  colnames(m) <- c(1, rep("", 23),25,rep("", 24),50,rep("", 24),75,rep("", 24),100,rep("", 24),125,rep("", 24),150,rep("", 24),175,rep("", 24),200)
  m <- m[,1:164] #BC is 15nt and L3 is 21nt
  column_ha = HeatmapAnnotation(
    colSum = anno_barplot(log2(colSums(m) + 1)),
    annotation_label=c(""),
    annotation_name_gp = gpar(fontsize = 7),
    height=unit(1, "cm"))

  G4_containing <- construct_info.df$type[match(rownames(m), construct_info.df$Construct.ID)]
  G4_containing[grep("pG4", G4_containing, fixed=T)] <- "Yes"
  G4_containing[grep("BS", G4_containing, fixed=T)] <- "No"
  row_ha = rowAnnotation(pG4 = G4_containing, col=list(pG4=c("Yes" = "black", "No" = "white")),
                              show_annotation_name=T, annotation_name_gp = gpar(fontsize = 6))
  
  set.seed(123)
  hm <- Heatmap(m, cluster_rows=F, cluster_columns=F, col=hm_col, top_annotation = column_ha, right_annotation = row_ha,
                show_row_names = F, use_raster = TRUE, raster_quality = 10, raster_by_magick = F,
                column_names_gp = gpar(fontsize = 7))
  
  return(hm)
})

ht_list = heatmaps.l[[1]] + heatmaps.l[[2]] + heatmaps.l[[3]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "GTP - KCl", show_heatmap_legend = F)

ht_list = heatmaps.l[[4]] + heatmaps.l[[5]] + heatmaps.l[[6]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "GTP - NaCl", show_heatmap_legend = F)

ht_list = heatmaps.l[[7]] + heatmaps.l[[8]] + heatmaps.l[[9]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "7dGTP - KCl", show_heatmap_legend = F)

ht_list = heatmaps.l[[10]] + heatmaps.l[[11]] + heatmaps.l[[12]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "7dGTP - NaCl", show_heatmap_legend = F)

```

```{r heatmap_legend, fig.width=12, fig.height=1}
draw(Legend(col_fun = hm_col, title = "RCPM", direction = "horizontal", grid_width = unit(8, "cm")))
```

\FloatBarrier
\pagebreak

To reduce the effect of too many rows on the visualization I determined the 500 constructs with the highest median RCPM across
the 12 experiments and used only these for the Heatmaps.

```{r filtered_metaprofile_heatmaps, fig.width=12, fig.height=3}

#hm_col <- circlize::colorRamp2(c(0, 5, 10, 20), c(viridis(4)[1], viridis(4)[2], viridis(4)[3], viridis(4)[4]))

hm_col <- circlize::colorRamp2(c(0, 10, 20, 40), c(viridis(4)[1], viridis(4)[2], viridis(4)[3], viridis(4)[4]))

top500 <- rowMedians(RCPMs_per_construct.df %>% as.matrix) %>% order(., decreasing=T)
top500 <- top500[1:500]

heatmaps.l <- lapply(RCPMs.l, function(l){
  m <- l %>% as.matrix
  rownames(m) <- l %>% names
  
  m <- m[top500,]
  
  row_order <- m[,19:200] %>% apply(1, which.max) %>% unlist  %>% order %>% rev
  m <- m[row_order,]

  colnames(m) <- c(1, rep("", 23),25,rep("", 24),50,rep("", 24),75,rep("", 24),100,rep("", 24),125,rep("", 24),150,rep("", 24),175,rep("", 24),200)
  m <- m[,1:164] #BC is 15nt and L3 is 21nt
  column_ha = HeatmapAnnotation(
    colSum = anno_barplot(log2(colSums(m) + 1)),
    annotation_label=c(""),
    annotation_name_gp = gpar(fontsize = 7),
    height=unit(1, "cm"))

  G4_containing <- construct_info.df$type[match(rownames(m), construct_info.df$Construct.ID)]
  G4_containing[grep("pG4", G4_containing, fixed=T)] <- "Yes"
  G4_containing[grep("BS", G4_containing, fixed=T)] <- "No"
  row_ha = rowAnnotation(pG4 = G4_containing, col=list(pG4=c("Yes" = "black", "No" = "white")),
                              show_annotation_name=T, annotation_name_gp = gpar(fontsize = 6))
  
  set.seed(123)
  hm <- Heatmap(m, cluster_rows=F, cluster_columns=F, col=hm_col, top_annotation = column_ha, right_annotation = row_ha,
                show_row_names = F, use_raster = TRUE, raster_quality = 10, raster_by_magick = F,
                column_names_gp = gpar(fontsize = 7))
  
  return(hm)
})

ht_list = heatmaps.l[[1]] + heatmaps.l[[2]] + heatmaps.l[[3]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "GTP - KCl", show_heatmap_legend = F)

ht_list = heatmaps.l[[4]] + heatmaps.l[[5]] + heatmaps.l[[6]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "GTP - NaCl", show_heatmap_legend = F)

ht_list = heatmaps.l[[7]] + heatmaps.l[[8]] + heatmaps.l[[9]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "7dGTP - KCl", show_heatmap_legend = F)

ht_list = heatmaps.l[[10]] + heatmaps.l[[11]] + heatmaps.l[[12]]
draw(ht_list, annotation_legend_side = "top", heatmap_legend_side = "bottom", ht_gap = unit(1, "cm"), column_title = "7dGTP - NaCl", show_heatmap_legend = F)

```

```{r filtered_heatmap_legend, fig.width=12, fig.height=1}
draw(Legend(col_fun = hm_col, title = "RCPM", direction = "horizontal", grid_width = unit(8, "cm")))
```

\FloatBarrier
\pagebreak


# Construct expression strength across experiments

Next, it was of interest if the "expression strength" of the constructs is the same across the 12 experiments or if there is an
experimental bias. For this purpose, the log2-transformed RCPM values of the constructs were used for pairwise correlations
among the 12 experiments. The correlations are presented in the heatmap.

```{r experiment_correlation_heatmap, fig.width = 8, fig.height=7}

#Prepare correlation matrix
corr.m <- sapply(1:12, function(i){
  sapply(1:12, function(j){
    cor(
      log2_RCPMs_per_construct.df[,i],
      log2_RCPMs_per_construct.df[,j]
      )
  })
})
colnames(corr.m) <- colnames(log2_RCPMs_per_construct.df)
rownames(corr.m) <- colnames(log2_RCPMs_per_construct.df)

#Prepare column annotation
column_ha = HeatmapAnnotation(
  base = rep(c("GTP", "7dGTP"), each=6),
  cation = rep(rep(c("KCl", "NaCl"), each=3),2),
  #replicate = rep(c(1,2,3), 4) %>% factor(., levels=c("1", "2", "3")),
  col=list(base = c("GTP" = "red", "7dGTP" = "orange"),
           cation=c("KCl" = "#1b8904", "NaCl" = "#0a1fb3")),
  show_annotation_name=T,
  annotation_legend_param = list(nrow=1))

#Create Heatmap
set.seed(123)
hm <- Heatmap(corr.m, cluster_rows = T, cluster_columns = T, col = viridis::viridis(200), show_column_names = T, show_row_dend = F, show_column_dend = T, border=T, top_annotation = column_ha,
        row_names_gp = gpar(fontsize = 9), column_names_gp = gpar(fontsize = 9), heatmap_legend_param = list(title="Pearson correlation", legend_direction = "horizontal"))

draw(hm, annotation_legend_side = "top", heatmap_legend_side = "bottom")
```

\FloatBarrier

As we can see the experiments cluster first by the used Guanine (GTP or 7dGTP) and second by the used cation (KCl or NaCl).

\pagebreak 

For more detailed insights the underlying correlations are shown in a ggpairs plot.

```{r experiment_correlation_ggpairs, fig.width = 12, fig.height=12}
#Create user-defined function
my_fn <- function(data, mapping, ...){
  ggplot(data, mapping) + rasterise(geom_pointdensity(size=.5), dpi=300) + scale_color_viridis() + geom_abline(intercept = 0, slope=1, col="red", size=.5) + lims(x=c(0,12), y=c(0,12))
}

#Create plot
ggpairs(log2_RCPMs_per_construct.df, lower = list(continuous = wrap(my_fn))
        ) + ggplot.theme.no.legend + theme(strip.text = element_text(size = 6))
```

\FloatBarrier
\pagebreak

After observing the clustering of experiments by Guanine and cation it was of interest to visualize the expression strength of
the individual constructs across the experiments. For this purpose, I first filtered for constructs with a median RCPM value
$\ge$ 10, followed by a z-score transformation by row.

```{r construct_variability_heatmap, fig.width = 8, fig.height=7}
RCPMs_per_construct_zscores.m <- RCPMs_per_construct.df %>% as.matrix
RCPMs_per_construct_zscores.m <- RCPMs_per_construct_zscores.m[RCPMs_per_construct_zscores.m %>% rowMedians >= 10,]
RCPMs_per_construct_zscores.m <- (RCPMs_per_construct_zscores.m  - (RCPMs_per_construct_zscores.m %>% rowMeans)) / rowSds(RCPMs_per_construct_zscores.m)


column_ha = HeatmapAnnotation(
  base = rep(c("GTP", "7dGTP"), each=6),
  cation = rep(rep(c("KCl", "NaCl"), each=3),2),
  #replicate = rep(c(1,2,3), 4) %>% factor(., levels=c("1", "2", "3")),
  col=list(base = c("GTP" = "red", "7dGTP" = "orange"),
           cation=c("KCl" = "#1b8904", "NaCl" = "#0a1fb3")),
  show_annotation_name=T,
  annotation_legend_param = list(nrow=1))

hm_col = circlize::colorRamp2(c(-3, -1, 0, 1, 3), c(viridis(5)[1], viridis(5)[2], viridis(5)[3], viridis(5)[4], viridis(5)[5]))
set.seed(123)
hm <- Heatmap(RCPMs_per_construct_zscores.m, cluster_rows = T, cluster_columns = T, col = hm_col, show_column_names = T, show_row_names = F, show_row_dend = F, show_column_dend = T, border=T,
              top_annotation = column_ha, row_names_gp = gpar(fontsize = 9), column_names_gp = gpar(fontsize = 9), heatmap_legend_param = list(title="z-score(RCPM)", legend_direction = "horizontal"),
              use_raster = T, raster_quality = 10, raster_by_magick=F,)

draw(hm, annotation_legend_side = "top", heatmap_legend_side = "bottom")
```

As we can see there are two main clusters of constructs, those with higher RCPM values in the GTP experiments and those with
higher RCPM values in the 7dGTP experiments. There is also a separation of KCl and NaCl but this not clearly evident in the
heatmap. A slight difference between KCl and NaCl becomes apparent in the GTP experiments.

\FloatBarrier
\pagebreak

# Correlation of technical construct replicates

In the next step it was of interest to see whether the expression strength of the technical replicates (constructs differing
only by barcode) was consistent across the experiments. For this purpose, pairwise correlations among all quartets of constructs
were done. The distribution of the mean of the 6 correlation coefficients is shown.

```{r correlation_technical_construct_replicates, fig.width=4, fig.height=3}

tech_reps.l <- split(log2_RCPMs_per_construct.df, construct_info.df$Region.ID)

tech_reps_mean_r.df <-lapply(tech_reps.l, function(df){
  r1 <- cor(df[1,] %>% as.numeric, df[2,] %>% as.numeric,)
  r2 <- cor(df[1,] %>% as.numeric, df[3,] %>% as.numeric,)
  r3 <- cor(df[1,] %>% as.numeric, df[4,] %>% as.numeric,)
  r4 <- cor(df[2,] %>% as.numeric, df[3,] %>% as.numeric,)
  r5 <- cor(df[2,] %>% as.numeric, df[4,] %>% as.numeric,)
  r6 <- cor(df[3,] %>% as.numeric, df[4,] %>% as.numeric,)
  
  mean_r <- mean(c(r1, r2, r3, r4, r5, r6))
  
  return(mean_r)
  
}) %>% unlist %>% as.data.frame %>% dplyr::rename(., mean_r = .) 


tech_reps_mean_r.df %>%
  ggplot(., aes(x=mean_r)) +
  geom_density() +
  ggplot.theme.no.legend +
  labs(x="Mean pearson correlation among the 4 technical replicates", y="Density")
```
As we can see, there are a lot of quartets showing consisten behaviours (close to 1.0). However, there are also quartets with no
or anticorrelation.

\FloatBarrier
\pagebreak

In the following, the pairwise correlations of the four quartets with the lowest mean correlation are shown. 

```{r anticorrelated_constructs, fig.height=5, fig.width=6}

mean_r_order <- order(tech_reps_mean_r.df$mean_r)

problematic_constructs.l <- lapply(mean_r_order[1:4], function(i){
  df <- tech_reps.l[[i]] %>% t %>% as.data.frame %>% add_rownames()
  colnames(df) <- c("exp", paste0("techrep", 1:4))
  
  ggarrange(
    ggplot(df, aes(x=techrep1, y=techrep2, fill=exp)) +
      geom_point(pch=21, size=3) +
      labs(x="log2(RCPM) tech. rep. 1", y="log2(RCPM) tech. rep. 2") +
      ggplot.theme.no.legend +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep1, y=techrep3, fill=exp)) +
      geom_point(pch=21, size=3) +
      labs(x="log2(RCPM) tech. rep. 1", y="log2(RCPM) tech. rep. 3") +
      ggplot.theme +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep1, y=techrep4, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 1", y="log2(RCPM) tech. rep. 4") +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep2, y=techrep3, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 2", y="log2(RCPM) tech. rep. 3") +
      scale_fill_manual(values=experiment_cols),
      
    ggplot(df, aes(x=techrep2, y=techrep4, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 2", y="log2(RCPM) tech. rep. 4") +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep3, y=techrep4, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 3", y="log2(RCPM) tech. rep. 4") +
      scale_fill_manual(values=experiment_cols),
    
    ncol=3, nrow=3, heights = c(10, 10, 1))
})

```

\FloatBarrier
\pagebreak

As a comparison, the pairwise correlations of the four quartets with the highest mean correlation are shown.

```{r correlated_constructs, fig.height=5, fig.width=6}

mean_r_order <- order(tech_reps_mean_r.df$mean_r, decreasing = T)

problematic_constructs.l <- lapply(mean_r_order[1:4], function(i){
  df <- tech_reps.l[[i]] %>% t %>% as.data.frame %>% add_rownames()
  colnames(df) <- c("exp", paste0("techrep", 1:4))
  
  ggarrange(
    ggplot(df, aes(x=techrep1, y=techrep2, fill=exp)) +
      geom_point(pch=21, size=3) +
      labs(x="log2(RCPM) tech. rep. 1", y="log2(RCPM) tech. rep. 2") +
      ggplot.theme.no.legend +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep1, y=techrep3, fill=exp)) +
      geom_point(pch=21, size=3) +
      labs(x="log2(RCPM) tech. rep. 1", y="log2(RCPM) tech. rep. 3") +
      ggplot.theme +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep1, y=techrep4, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 1", y="log2(RCPM) tech. rep. 4") +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep2, y=techrep3, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 2", y="log2(RCPM) tech. rep. 3") +
      scale_fill_manual(values=experiment_cols),
      
    ggplot(df, aes(x=techrep2, y=techrep4, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 2", y="log2(RCPM) tech. rep. 4") +
      scale_fill_manual(values=experiment_cols),
    
    ggplot(df, aes(x=techrep3, y=techrep4, fill=exp)) +
      geom_point(pch=21, size=3) +
      ggplot.theme.no.legend +
      labs(x="log2(RCPM) tech. rep. 3", y="log2(RCPM) tech. rep. 4") +
      scale_fill_manual(values=experiment_cols),
    
    ncol=3, nrow=3, heights = c(10, 10, 1))
})




```


\FloatBarrier
\pagebreak

# Definition of read-through and RTstop regions.

For a first quantitative overview I defined two regions in each construct, namely the read through region ranging from position
1 to 18 and the RTstop region ranging from position 51 to 100. Afterwards, I summed up the RCPM values in each region and calculate
log2(RTstop/read through).

```{r define_regions_and_calc_log2FC}

RT_stop_enrichment.df <- lapply(RCPMs.l, function(l){
  m <- l %>% as.matrix

  a <- (m[,51:100] %>% rowSums)+1
  b <- (m[,1:18] %>% rowSums)+1
  return(log2(a/b))
}) %>% do.call("cbind", .) %>% as.data.frame

rownames(RT_stop_enrichment.df) <- rownames(RCPMs_per_construct.df)

```

For a first overview I create MA plots to monitor the RTstop enrichment relative to the expression strength of the underlying
construct.

```{r RT_stop_enrichment_MA, fig.width=8, fig.height=7}
MA.df <- RT_stop_enrichment.df %>% add_rownames() %>% pivot_longer(., cols=2:13, names_to = "experiment", values_to = "log2FC") %>%
  mutate(experiment = factor(experiment, levels = names(experiment_cols))) 

MA.df$expression <- MA.df %>% apply(., 1, function(row){
  log2_RCPMs_per_construct.df[match(row["rowname"], rownames(log2_RCPMs_per_construct.df)), row["experiment"]]
}) 
  
MA.df %>% 
  ggplot(., aes(x=expression, y=log2FC)) +
  rasterise(geom_pointdensity(), dpi=300) +
  scale_color_viridis() + 
  labs(x="log2(RCPM for construct)", y="log2FC(RTstops / read through)") +
  facet_wrap(~experiment, ncol=3, nrow=4) + ggplot.theme.no.legend
```

\FloatBarrier

Not surprising the log2FC values are capped by the underlying construct expression strength up to a log2(RCPM) of about 6.

\pagebreak

To account for the dependence on the construct expression strength, I limited the constructs to those with a median RCPM value of 
$\ge$ 64 (2^6) and updated the MA plots.

```{r RT_stop_enrichment_MA_filtered, fig.width=8, fig.height=7}

filtered_RT_stop_enrichment.df <- RT_stop_enrichment.df[RCPMs_per_construct.df %>% as.matrix %>% rowMedians >= 64, ]

MA.df <- filtered_RT_stop_enrichment.df %>% add_rownames() %>% pivot_longer(., cols=2:13, names_to = "experiment", values_to = "log2FC") %>%
  mutate(experiment = factor(experiment, levels = names(experiment_cols))) 

MA.df$expression <- MA.df %>% apply(., 1, function(row){
  log2_RCPMs_per_construct.df[match(row["rowname"], rownames(log2_RCPMs_per_construct.df)), row["experiment"]]
}) 
  
MA.df %>% 
  ggplot(., aes(x=expression, y=log2FC)) +
  rasterise(geom_pointdensity(), dpi=300) + 
  scale_color_viridis() + 
  labs(x="log2(RCPM for construct)", y="log2FC(RTstops / read through)") +
  facet_wrap(~experiment, ncol=3, nrow=4) + ggplot.theme.no.legend
```

\FloatBarrier

There are still some constructs with lower expression strength in individual experiments, but the bias of such
lowly expressed constructs should strongly reduced.

\pagebreak

For a more detailed overview I created a density distribution for the log2FCs. The distributions in the front are the filtered
ones (only constructs with median RCPM value $\ge$ 64). The distribution in the back is without filtering (all 12.000 constructs).

```{r RT_stop_enrichment_distributions, fig.width=8, fig.height=7}

rbind(
  RT_stop_enrichment.df %>% pivot_longer(., cols=everything(), names_to = "experiment", values_to = "log2FC") %>%
    mutate(experiment = factor(experiment, levels = names(experiment_cols))) %>% 
    mutate(filtered=F), 

  filtered_RT_stop_enrichment.df %>% pivot_longer(., cols=everything(), names_to = "experiment", values_to = "log2FC") %>%
    mutate(experiment = factor(experiment, levels = names(experiment_cols))) %>% 
    mutate(filtered=T)
) %>% 
  ggplot(., aes(x=log2FC, fill=experiment, alpha = filtered)) +
  geom_density() + 
  scale_fill_manual(values=experiment_cols) +
  scale_alpha_manual(values=c("TRUE"=.75, "FALSE"=.25)) +
  geom_vline(xintercept=0, linetype="dashed") +
  labs(x="log2FC(RTstops / read through)", y="Density") +
  facet_wrap(~experiment, ncol=3, nrow=4) + ggplot.theme.no.legend



```

As we can see the removal of constructs with weak expression strength has a strong effect and leads to more bell-shaped distributions.
We can also see higher density in the log2FC region > 0 for the GTP KCl experiments than for the other experiments.

# Next steps

The next step will be to find a good "cutpoint" in the distributions to discriminate between constructs with an enrichment of RTstops
and constructs with no enrichment. A simple cutoff at 0 does not make sense as we would loose for instance the construct containing the
G4 in the downstream intron of MYL6 (see Construct.ID 10119 in the Shiny app). It is a clear G4, which we should keep. 

An alternative to the cutpoint could be to calculate z-scores for the log2FCs across the experiments.


# Session Information

```{r}
sessionInfo()
```
