---
title: "HNRNPH RTstop sites - classification"
author: "Mirko Brueggemann"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    keep_tex: yes
  html_document:
    df_print: paged
  pdf_document: default
graphics: yes
geometry: margin=1in
fontsize: 11pt
header-includes:
- \usepackage{placeins}
---

```{r knitr off, echo=FALSE, message=FALSE,cache=FALSE, results="hide", render=FALSE ,warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE,message=FALSE)
```


# Analysis Description

This report shows how to classify G4 peak containing constructs based on the number and positions of peaks inside the construct. The KCl RT-stop signal from all 3 replicates is summed up and represented as a matrix. Counts in this matrix are min-max normalized and spline transformed to smooth out the signal. 

```{r,loadLibs}
### ============================================================================
### Load Packages
### ----------------------------------------------------------------------------
### 
library(rtracklayer)
library(GenomicRanges)
library(ggplot2)
library(AnnotationDbi)
library(dplyr)
library(reshape2)
library(UpSetR)
library(GenomicFeatures)
library(kableExtra)
library(knitr)
library(ggrepel)
library(gridExtra)
library(grid)
library(viridis)
library(BindingSiteFinder)
library(ComplexHeatmap)
library(forcats)
library(ggtext)
library(patchwork)
library(ggforce)
library(ggpointdensity)
library(knitr)
library(tidyr)
library(ggpointdensity)
library(ggsci)
library(DESeq2)
library(edgeR)
library(ggrastr)
```

```{r}
myGrey = "#808080"
# Define plotting theme
theme_nice <- function(size = 12) {
    baseSize = size
    largeSize = size + 2
    smallSize = size - 2
    theme_bw() +
        theme(axis.text = element_text(size = smallSize),
              axis.title = element_text(size = baseSize),
              strip.text.x = element_text(size = baseSize),
              ## legend
              legend.text = element_text(size = baseSize),
              legend.title = element_text(size = baseSize),
              ## title
              title = element_text(size = largeSize),
              plot.title = element_text(size = largeSize),
              plot.subtitle = element_text(size = smallSize),
              plot.title.position = 'plot',
              ## caption
              plot.caption = element_text(hjust = 0,
                                          face= "italic", size = smallSize),
              plot.caption.position = "plot",
        )
}

theme_nice_pie <- function(size = 12) {
    baseSize = size
    largeSize = size + 2
    smallSize = size - 2
    theme_void() +
        theme(## legend
              legend.text = element_text(size = baseSize),
              legend.title = element_text(size = baseSize),
              ## title
              title = element_text(size = largeSize),
              plot.title = element_text(size = largeSize),
              plot.subtitle = element_text(size = smallSize),
              plot.title.position = 'plot',
              ## caption
              plot.caption = element_text(hjust = 0,
                                          face= "italic", size = smallSize),
              plot.caption.position = "plot",
        )
}

theme_pub <- function(){
  theme_bw() + 
  theme(axis.text = element_text(size = 6),
        axis.title = element_text(size = 8),
        title = element_text(size = 10),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        plot.title = element_text(size = 10),
        strip.text.x = element_text(size = 8))
} 

theme_md <- function() {
  theme(              ## nicer text formatting
              plot.title = ggtext::element_markdown(),
              plot.caption = ggtext::element_markdown(),
              plot.subtitle = ggtext::element_markdown(),
              axis.title.x = ggtext::element_markdown(),
              axis.title.y = ggtext::element_markdown())
}
```

```{r}
minMaxNorm <- function(m) {
  nM = apply(m, 1, function(x){
    n = ((x - min(x)) / (max(x) - min(x)))
    return(n)
  })
  nM = t(nM)
  return(nM)
}

smoothing <- function(y, lambda, dim){
  p <- length(y)
  
  which(y != max(y))
  if ( length(which(y != max(y))) != 0 ){
    x = c(1:p)
    y.smooth = smooth.spline(x, y, spar = lambda)
    x_new = seq(1, p, p/dim)
    y_new = predict(y.smooth, x_new)$y
    y_new = y_new / max(y_new)
    return(y_new)
  } 
  
  x_new <- seq(1, p, p / dim)
  y_new <- rep( max(y), length(x_new) )
  return(y_new)
}
```


# Count smoothing and transformation

```{r}
# make signal matrix
constructSignal = readRDS("../rds/RTstop.profiling.l.rds")
masterMatrKCL = as.matrix(constructSignal$GTP_KCL_rep1) + as.matrix(constructSignal$GTP_KCL_rep2) + as.matrix(constructSignal$GTP_KCL_rep3)
rownames(masterMatrKCL) = names(constructSignal$GTP_KCL_rep1)

# load G4 peaks
G4Peaks = readRDS("../03_PeakDifferencesFinal/data/G4Peaks.rds")
masterMatrKCL = masterMatrKCL[rownames(masterMatrKCL) %in% G4Peaks$constructID,]

# normalize matrix
normKCL = minMaxNorm(masterMatrKCL)

# apply different smoothing levels
smoothKCL_par1 = t(apply(normKCL, 1, smoothing, lambda=0.4, dim=400))
smoothKCL_par2 = t(apply(normKCL, 1, smoothing, lambda=0.5, dim=400))
smoothKCL_par3 = t(apply(normKCL, 1, smoothing, lambda=0.6, dim=400))
saveRDS(smoothKCL_par1, file = "/Users/mirko/Projects/HNRNPH/02_markdowns/RTstop-profiling/04_PeakClassification/data/smoothKCL_par1.rds")
saveRDS(smoothKCL_par2, file = "/Users/mirko/Projects/HNRNPH/02_markdowns/RTstop-profiling/04_PeakClassification/data/smoothKCL_par2.rds")
saveRDS(smoothKCL_par3, file = "/Users/mirko/Projects/HNRNPH/02_markdowns/RTstop-profiling/04_PeakClassification/data/smoothKCL_par3.rds")
```

```{r, fig.width=8, fig.height=5, fig.cap="\\textbf{Heatmaps of different smooting levels before clustering.}"}
set.seed(1234)
custom.col = viridis(10, option = "G", direction = -1)

idx = sample(1:(nrow(smoothKCL_par1)), 500)

m1 = smoothKCL_par1[idx,]
m2 = smoothKCL_par2[idx,]
m3 = smoothKCL_par3[idx,]

h1 = Heatmap(m1, column_title = "Norm+Smooth 0.4", name = "Xlinks",
             cluster_columns = F, cluster_rows = T,
             show_row_names = FALSE,
             show_column_names = T, border = T, col = custom.col,
             heatmap_legend_param = list(
               at = c(0,0.5, 1), labels = c("0", "0.5", "1"),
               legend_width = unit(6, "cm"),
               title_position = "topleft", direction = "horizontal"
             ), use_raster = TRUE
)
h2 = Heatmap(m2, column_title = "Norm+Smooth 0.5", name = "Xlinks",
             cluster_columns = F, cluster_rows = T,
             show_row_names = FALSE,
             show_column_names = T, border = T, col = custom.col,
             heatmap_legend_param = list(
               at = c(0,0.5, 1), labels = c("0", "0.5", "1"),
               legend_width = unit(6, "cm"),
               title_position = "topleft", direction = "horizontal"
             ), use_raster = TRUE
)
h3 = Heatmap(m3, column_title = "Norm+Smooth 0.6", name = "Xlinks",
             cluster_columns = F, cluster_rows = T,
             show_row_names = FALSE,
             show_column_names = T, border = T, col = custom.col,
             heatmap_legend_param = list(
               at = c(0,0.5, 1), labels = c("0", "0.5", "1"),
               legend_width = unit(6, "cm"),
               title_position = "topleft", direction = "horizontal"
             ), use_raster = TRUE
)

l = h1 + h2 + h3
draw(l, heatmap_legend_side = "bottom")
```

```{r, laodUMAP}
# load UMAP classification results
umapKCL_par1 = readRDS("/Users/mirko/Projects/HNRNPH/02_markdowns/RTstop-profiling/04_PeakClassification/data/umapKCL_par1.rds")
umapKCL_par2 = readRDS("/Users/mirko/Projects/HNRNPH/02_markdowns/RTstop-profiling/04_PeakClassification/data/umapKCL_par2.rds")
umapKCL_par3 = readRDS("/Users/mirko/Projects/HNRNPH/02_markdowns/RTstop-profiling/04_PeakClassification/data/umapKCL_par3.rds")
```

```{r, fig.width=10,fig.height=4, fig.cap="\\textbf{UMAP transformation with Kmeans clustering of different smooting settings.}"}
group_par1 = kmeans(umapKCL_par1$layout, centers = 2)
df1 = data.frame(x = umapKCL_par1$layout[,1],
                y = umapKCL_par1$layout[,2],
                group = factor(group_par1$cluster))

group_par2 = kmeans(umapKCL_par2$layout, centers = 2)
df2 = data.frame(x = umapKCL_par2$layout[,1],
                y = umapKCL_par2$layout[,2],
                group = factor(group_par2$cluster))

group_par3 = kmeans(umapKCL_par3$layout, centers = 2)
df3 = data.frame(x = umapKCL_par3$layout[,1],
                y = umapKCL_par3$layout[,2],
                group = factor(group_par3$cluster))

p1 = ggplot(df1, aes(x, y, color = group)) +
  ggrastr::geom_point_rast(size = 0.2) +
  theme_nice() +
  scale_color_manual(values = viridis(n = 5, option = "mako")[2:4]) +
  guides(colour = guide_legend(override.aes = list(size=10))) +
  theme(legend.position = "top") +
  labs(title = "Smooth 0.4")

p2 = ggplot(df2, aes(x, y, color = group)) +
  ggrastr::geom_point_rast(size = 0.2) +
  theme_nice() +
  scale_color_manual(values = viridis(n = 5, option = "mako")[2:4]) +
  guides(colour = guide_legend(override.aes = list(size=10))) +
  theme(legend.position = "top") +
  labs(title = "Smooth 0.5") 

p3 = ggplot(df3, aes(x, y, color = group)) +
  ggrastr::geom_point_rast(size = 0.2) +
  theme_nice() +
  scale_color_manual(values = viridis(n = 5, option = "mako")[2:4]) +
  guides(colour = guide_legend(override.aes = list(size=10))) +
  theme(legend.position = "top") +
  labs(title = "Smooth 0.6")

p1 + p2 + p3
```


NOTE:

* higher smoothing leads to more distinct groups
* I selected for smoothing level 0.6
* Smoothing level 0.5 might also be fine
* The smoothing defines how good the signal can be seen -> and then later classified
 * Larger smoothing leads to less accurate peaks, but constructs can be easier clustered by their similarities/ differences
 * Smaller smoothing leads to very accurate peaks, but differences between constructs are hard to summarize/cluster, because more variation exists

\FloatBarrier
\pagebreak

# Kmeans and density-based clustering

```{r, fig.width=10, fig.height=4, fig.cap="\\textbf{Kmeans clustering with 4 centroids.}"}
set.seed(1234)
group_par1 = kmeans(umapKCL_par3$layout, centers = 4)
df1 = data.frame(x = umapKCL_par3$layout[,1],
                y = umapKCL_par3$layout[,2],
                group = factor(group_par1$cluster))

p1 = ggplot(df1, aes(x, y, color = group)) +
  geom_point(size = 0.2) +
  theme_nice() +
  scale_color_manual(values = viridis(n = 6, option = "mako")[2:8]) +
  guides(colour = guide_legend(override.aes = list(size=10), nrow = 1)) +
  theme(legend.position = "top") +
  labs(title = "Smooth 0.6") 

p2 = ggplot(df1, aes(x, y)) +
  geom_pointdensity(size = 0.2) +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top") +
  labs(title = "Smooth 0.6") +
    guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(20, 'lines'), barheight = unit(.5, 'lines'))) 
p1 + p2
```

```{r, fig.width=6, fig.height=4, fig.cap="\\textbf{K nearest neighbors plot with k=35.} K equals the MinPts option in the dbscan. Line at 0.6 represents eps param."}
library(fpc)
library(dbscan)
set.seed(1234)
dbscan::kNNdistplot(umapKCL_par3$layout, k = 35)
abline(h = 0.6, lty = 2)
```

```{r, fig.width=12, fig.height=6, fig.cap="\\textbf{UMAP transformation with dbscan clustering; eps=0.6, MinPts=35.}"}
group_par1 = dbscan::dbscan(umapKCL_par3$layout, eps = 0.6, minPts = 35)

df1 = data.frame(x = umapKCL_par3$layout[,1],
                 y = umapKCL_par3$layout[,2],
                 group = factor(group_par1$cluster))

p1 = ggplot(df1, aes(x, y, fill = group)) +
  geom_point(shape = 21) +
  theme_nice() +
  scale_fill_viridis(option = "mako", discrete = T, direction = -1) +
  guides(colour = guide_legend(override.aes = list(size=1), nrow = 3)) +
  theme(legend.position = "top", aspect.ratio = 1)

p2 = ggplot(df1, aes(x, y)) +
  geom_pointdensity(size = 0.8) +
  # ggrastr::rasterise(geom_pointdensity(size = 0.2)) +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top", aspect.ratio = 1) +
  guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(10, 'lines'), barheight = unit(.5, 'lines'))) 

p3 = ggplot(subset(df1, group != 0), aes(x, y, fill = group)) +
  geom_point(shape = 21) +
  theme_nice() +
  scale_fill_viridis(option = "mako", discrete = T, direction = -1) +
  guides(colour = guide_legend(override.aes = list(size=1), nrow = 3)) +
  theme(legend.position = "top", aspect.ratio = 1)

p1 + p2 + p3 + patchwork::plot_layout(guides = "collect") & theme(legend.position = "top")
```

```{r, fig.width=10, fig.height=10, fig.cap="\\textbf{Smoothed crosslink profiles split by density based clustering.}"}
set.seed(1234)

s = split.data.frame(smoothKCL_par3, group_par1$cluster)
tmp = lapply(1:length(s), function(x){
  z = floor(ncol(s[[1]]) / 2)
  d = data.frame(Group = factor(rep(x, ncol(s[[x]]))),
                 value = colMeans(s[[x]], na.rm = TRUE),
                 pos = -z:z)
  
})
df = dplyr::bind_rows(tmp, .id = "variable")

ggplot(df, aes(x = pos, y = value, group = Group)) +
  geom_line(size = 1) +
  scale_color_viridis(option = "mako", discrete = T, direction = -1) +
  theme_nice() +
  theme(legend.position = "none") +
  # guides(colour = guide_legend(override.aes = list(size=10), nrow = 1))  +
  facet_wrap(~Group)

```

NOTE:

* density based clustering works better than kmeans clustering
* cluster 0 of the desnsity based clustering holds all outliers, which is why no clear peak can be seen outside of the read-through region
* As a result of the high smoothing the resolution of the classification does not allow an accurate classification by the number of peaks
 * less strong smoothing does solve the problem but leads to much smaller groups, since more variation is included


\FloatBarrier
\pagebreak

# Subselection on main groups

```{r, fig.width=8, fig.height=4, fig.cap="\\textbf{Number of constructs per cluster.}"}
nClusterCutoff = 100

df = table(group_par1$cluster) %>% 
  as.data.frame()
p1 = ggplot(df, aes(x = Var1, y = Freq)) +
  geom_col() + 
  coord_flip() +
  theme_nice() +
  theme(axis.text = element_text(size = 8) ) +
  geom_hline(yintercept = nClusterCutoff) +
  labs(
    y = "Count",
    x = "Cluster"
  )

groupSel = data.frame(group = group_par1$cluster) %>% 
  as.data.frame() %>%
  group_by(group) %>%
  summarize(n = n())

groupSel = groupSel$group[groupSel$n > nClusterCutoff]

p2 = ggplot(df, aes(x = reorder(Var1, -Freq), y = Freq, group = 1)) +
  geom_point() +
  geom_line() +
  theme_nice() +
  geom_hline(yintercept = nClusterCutoff) +
  labs(x = "Cluster", y = "Count") +
  theme(axis.text = element_text(size = 8) ) 

p1 + p2 
```

```{r, fig.width=10, fig.height=5, fig.cap="\\textbf{UMAP transformation with dbscan clustering on main groups.}"}
df = data.frame(x = umapKCL_par3$layout[,1],
                y = umapKCL_par3$layout[,2],
                group = factor(group_par1$cluster))
df = df[df$group != 0,]
df1 = df[df$group %in% groupSel,]
df2 = df[!df$group %in% groupSel,]

p1 = ggplot() +
  geom_point(data = df2, aes(x, y), fill = "white", shape = 21) +
  geom_point(data = df1, aes(x, y, fill = group), shape = 21) +
  theme_nice() +
  scale_fill_npg() +
  # scale_fill_viridis(option = "mako", discrete = T, direction = -1) +
  guides(colour = guide_legend(override.aes = list(size=1), nrow = 3)) +
  theme(legend.position = "top", aspect.ratio = 1) 

p2 = ggplot(df1, aes(x, y)) +
  geom_pointdensity() +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top", aspect.ratio = 1) +
    guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(10, 'lines'), barheight = unit(.5, 'lines'))) 

p3 = ggplot() +
  geom_point(data = df1, aes(x, y, fill = group), shape = 21) +
  theme_nice() +
  scale_fill_npg() +
  # scale_fill_viridis(option = "mako", discrete = T, direction = -1) +
  guides(colour = guide_legend(override.aes = list(size=1), nrow = 3)) +
  theme(legend.position = "top", aspect.ratio = 1)

p1 + p2 + p3 + patchwork::plot_layout(guides = "collect") & theme(legend.position = "top")
```

```{r,fig.width=4, fig.height=6, fig.cap="\\textbf{Smoothed crosslink heatmaps split by main groups.}"}
set.seed(1234)
s = split.data.frame(smoothKCL_par3, group_par1$cluster)
s = s[names(s) %in% groupSel]
s = s[names(s) != 0]

# s = sapply(s, head, n = 500)

test = lapply(seq_along(s), function(x){
  h = Heatmap(s[[x]],
          row_title = names(s[x]), name = "Xlinks",
          cluster_rows = FALSE, cluster_columns = FALSE, 
          show_column_names = T, border = T, col = custom.col,
          use_raster = TRUE, show_row_names = FALSE,
          heatmap_legend_param = list(
               at = c(0,0.5, 1), labels = c("0", "0.5", "1"),
               legend_width = unit(6, "cm"),
               title_position = "topleft", direction = "horizontal"
             )
  )
  h
})

# l = test[[1]] %v% test[[2]] %v% test[[3]] %v% test[[4]] 
# l = test[[1]] %v% test[[2]] %v% test[[3]] %v% test[[4]] %v% test[[5]] %v% test[[6]] %v% test[[7]] 
l = test[[1]] %v% test[[2]] %v% test[[3]] %v% test[[4]] %v% test[[5]] %v% test[[6]] %v% test[[7]]  %v% test[[8]]  %v% test[[9]]  %v% test[[10]] 
draw(l, heatmap_legend_side = "bottom")
```

```{r, fig.width=10, fig.height=4, fig.cap="\\textbf{Smoothed crosslink profiles split by main groups.}"}
set.seed(1234)

s = split.data.frame(smoothKCL_par3, group_par1$cluster)
tmp = lapply(1:length(s), function(x){
  # z = floor(ncol(s[[1]]) / 2)
  z = ncol(s[[1]])
  d = data.frame(Group = factor(rep(x, ncol(s[[x]]))),
                 value = colMeans(s[[x]], na.rm = TRUE),
                 pos = 1:z)
  
})
df = dplyr::bind_rows(tmp, .id = "variable")
df = df[df$Group %in% groupSel,]
df = df[df$Group != 0,]

ggplot(df, aes(x = pos, y = value, color = Group, group = Group)) +
  geom_line(size = 1) +
  scale_color_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  facet_wrap(~Group, ncol = 5)

```

NOTE:

* Very few groups contain most of the constructs
* These main groups show a broad spread of shapes
 * Most different are groups 5,6 with a large peak at the read-through region and only small peaks in the rest of the construct
 * All other groups contain major peaks outside of the read-through part and are different in terms of the location of these peaks


\FloatBarrier
\pagebreak

# Example profiles for each main group

```{r, fig.width=10, fig.height=10, fig.cap="\\textbf{Example profiles for each main cluster - smoothed.}"}
set.seed(1234)
idx = sample(1:100, 6)

sSel = s[names(s) %in% groupSel]
sSel = sSel[names(sSel) != 0]
tmp = lapply(sSel, function(x) {
  # x = sSel$`1`
  xx = x[idx,] %>%
    as.data.frame() %>%
    tibble::rownames_to_column("constructID") %>%
    pivot_longer(-constructID) %>%
    mutate(pos = rep(1:399,6))
    # mutate(pos = rep(-199:199,6))
  return(xx)
})
df = dplyr::bind_rows(tmp, .id = "Group")
df = df %>% unite("BothLabels", Group, constructID, sep = "_", remove = FALSE)
df$BothLabels = factor(df$BothLabels)
x = levels(df$BothLabels)
df$BothLabels = factor(df$BothLabels, levels = x[order(nchar(x),x)])

ggplot(df, aes(x = pos, y = value, color = Group)) +
  geom_line(size = 1) +
  scale_color_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  facet_wrap(~BothLabels, ncol = 6) +
  theme(
  strip.background = element_blank(),
  # axis.text = element_text(size = 6)
  strip.text.x = element_text(size = 8)
)
```

```{r, fig.width=10, fig.height=10, fig.cap="\\textbf{Example profiles for each main cluster - normalized}"}
tmp = normKCL[rownames(normKCL) %in% unique(df$constructID),] %>%
  as.data.frame() %>%
  tibble::rownames_to_column("constructID") %>%
  pivot_longer(-constructID) 

idx = match(tmp$constructID, df$constructID)
tmp$Group = df$Group[idx]
tmp$BothLabels = df$BothLabels[idx]
# tmp$pos = rep(-99:100,6*(length(groupSel)-1))
tmp$pos = rep(1:200,6*(length(groupSel)-1))

ggplot(tmp, aes(x = pos, y = value, color = Group)) +
  geom_line(size = 1) +
  scale_color_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  facet_wrap(~BothLabels, ncol = 6) +
  theme(
  strip.background = element_blank(),
  # axis.text = element_text(size = 6),
  strip.text.x = element_text(size = 8)
)
```

```{r, fig.width=10, fig.height=10, fig.cap="\\textbf{Example profiles for each main cluster - raw}"}
tmp = masterMatrKCL[rownames(masterMatrKCL) %in% unique(df$constructID),] %>%
  as.data.frame() %>%
  tibble::rownames_to_column("constructID") %>%
  pivot_longer(-constructID) 

idx = match(tmp$constructID, df$constructID)
tmp$Group = df$Group[idx]
tmp$BothLabels = df$BothLabels[idx]
# tmp$pos = rep(-99:100,6*(length(groupSel)-1))
tmp$pos = rep(1:200,6*(length(groupSel)-1))

ggplot(tmp, aes(x = pos, y = value, color = Group)) +
  geom_line(size = 1) +
  scale_color_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  facet_wrap(~BothLabels, ncol = 6, scales = "free_y") +
  theme(
  strip.background = element_blank(),
  axis.text = element_text(size = 6),
  strip.text.x = element_text(size = 8)
)
```

NOTE:

* Min-max normalization does not alter the relative hight of the peaks
* Smoothing strongly influences how the signal is distributed of the construct
 * the read-through part for example get summarized to a ~50nt wide peak region
* But still the smoothed signal does follow the raw signal

\FloatBarrier
\pagebreak

# G4 strength on construts of main groups

```{r, fig.width=8, fig.height=5, fig.cap="\\textbf{UMAP transformation of clustered groups colored by KCL to total ratio (G4 peak strength).} The G4 strenght per construct is represented as the maximum of its G4 peaks."}
df = data.frame(x = umapKCL_par3$layout[,1],
                y = umapKCL_par3$layout[,2],
                group = factor(group_par1$cluster))
df = df[df$group != 0,]
df$constructID = rownames(df)



dfG4Peaks = G4Peaks %>%
  as.data.frame() %>%
  group_by(constructID) %>%
  summarize(max_GTP_KCL_ratio = max(GTP_KCL_ratio),
            max_GTP_NACL_ratio = max(GTP_NACL_ratio),
            GTP_KCL_total, max_log2FoldChange = max(log2FoldChange))

idx = match(df$constructID, dfG4Peaks$constructID)
df$max_GTP_KCL_ratio = dfG4Peaks$max_GTP_KCL_ratio[idx]
df$max_GTP_NACL_ratio = dfG4Peaks$max_GTP_NACL_ratio[idx]
df$GTP_KCL_total = dfG4Peaks$GTP_KCL_total[idx]
df$max_log2FoldChange = dfG4Peaks$max_log2FoldChange[idx]

q = quantile(df$max_GTP_KCL_ratio, seq(0,1, by = 0.1))
df$bins = as.factor(findInterval(df$max_GTP_KCL_ratio, q, all.inside = TRUE))

df1 = df[df$group %in% groupSel,]
df2 = df[!df$group %in% groupSel,]

p1 = ggplot() +
  geom_point(data = df, aes(x, y, color = max_GTP_KCL_ratio),  size = 0.8) +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top", aspect.ratio = 1) +
  guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(10, 'lines'), barheight = unit(.5, 'lines'))) +
  labs(title = "All groups")

p2 = ggplot() +
  geom_point(data = df1, aes(x, y, color = max_GTP_KCL_ratio), size = 0.8) +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top", aspect.ratio = 1) +
  guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(10, 'lines'), barheight = unit(.5, 'lines'))) +
  labs(title = "Main groups")

(p1 + p2) 
```

```{r, fig.width=8, fig.height=5, fig.cap="\\textbf{UMAP transformation of clustered groups colored by NACL to total ratio.} The ratio per construct is represented as the maximum of its G4 peaks."}
p1 = ggplot() +
  geom_point(data = df, aes(x, y, color = max_GTP_NACL_ratio),  size = 0.8) +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top", aspect.ratio = 1) +
    guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(10, 'lines'), barheight = unit(.5, 'lines'))) +
  labs(title = "All groups")

p2 = ggplot() +
  geom_point(data = df1, aes(x, y, color = max_GTP_NACL_ratio), size = 0.8) +
  theme_nice() +
  scale_color_viridis(option = "A") +
  theme(legend.position = "top", aspect.ratio = 1) +
    guides(color = guide_colorbar(title.position = 'top', title.hjust = 0.5,
                                barwidth = unit(10, 'lines'), barheight = unit(.5, 'lines'))) +
  labs(title = "Main groups")
(p1 + p2) 
```

```{r, fig.width=10, fig.height=4, fig.cap="\\textbf{G4 strength per cluster group.} Each box contains the indicated values for each construt. Boxes are separated by the indicated cluster group"}
library(forcats)
df1$group = droplevels.factor(df1$group)

p1 = ggplot(df1, aes(x = reorder(group,-max_GTP_KCL_ratio, na.rm = TRUE), y = max_GTP_KCL_ratio, fill = group)) +
  geom_boxplot() +
  scale_fill_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  labs(x = "Cluster group")

p2 = ggplot(df1, aes(x = reorder(group,-max_GTP_KCL_ratio, na.rm = TRUE), y = log2(GTP_KCL_total), fill = group)) +
  geom_boxplot() +
  scale_fill_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  labs(x = "Cluster group")

p3 = ggplot(df1, aes(x = reorder(group,-max_GTP_KCL_ratio, na.rm = TRUE), y = max_log2FoldChange, fill = group)) +
  geom_boxplot() +
  scale_fill_npg() +
  theme_nice() +
  theme(legend.position = "none") +
  labs(x = "Cluster group")

p1 + p2 + p3 
```

NOTE:

* Cluster groups 1,4,5 show by far the lowest KCL ratio -> consist of constructs with weak G4s
 * these groups show the majority of the signal at the read-through part of the construct
* Cluster group 5 shows the highest overall read coverage -> this is why these constructs have significant G4s.
 * small changes can be detected if the read signal/ statistical power is high
 * Yet they have the samllest log2 fold-change


\FloatBarrier
\pagebreak

# Session Information

```{r, session_info, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
```
