---
title: "Not predictable G4s"
author: "Mario Keller"
date: "`r format(Sys.time(), '%B %e, %Y')`"
geometry: "left=1cm,right=1cm,top=1.5cm,bottom=1.5cm"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    toc: true
    toc_depth: 3
    number_sections: true

header-includes:
- \makeatletter\renewcommand*{\fps@figure}{h}\makeatother
- \usepackage{placeins}
---

```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyr)
library(egg)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(GenomicRanges)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.pos = 'h', fig.align = "center", fig.width = 6, fig.height=6)
```

```{r ggplot.themes, echo=FALSE, message=FALSE, warning=FALSE}
ggplot.theme <- theme_bw() +
      theme(
        axis.text=element_text(size=7,face="bold", color = "black"),
        axis.title.x=element_text(size=8,face="bold"),
        axis.title.y=element_text(size=8,face="bold", vjust = 0),
        plot.title = element_text(size=8,face="bold"),
        legend.text = element_text(size=8,face="bold"),
        legend.title = element_text(size=8,face="bold"),
        legend.position = "bottom",
        legend.direction="horizontal")

ggplot.theme.no.legend <- theme_bw() +
      theme(
        axis.text=element_text(size=7,face="bold", color = "black"),
        axis.title.x=element_text(size=8,face="bold"),
        axis.title.y=element_text(size=8,face="bold", vjust = 0),
        plot.title = element_text(size=8,face="bold"),
        legend.text = element_text(size=8,face="bold"),
        legend.title = element_text(size=8,face="bold"),
        legend.position = "none",
        legend.direction="vertical")
```

# Objectives

The aim of this analysis was the identification of measured G4s (those with a positive and significant log2FC(KCl vs NaCl)) that were not predicted by pqsfinder or QGRS Mapper.

```{r loadAndPrepareData}
#Mirko's peaks and their regulation
load("/Users/mariokeller/projects/HNRNPH_project/RTstop-profiling/03_PeakDifferencesFinal/data/peaksChange.rds")

#Mirko's updated constructInfo dataframe
load("/Users/mariokeller/projects/HNRNPH_project/RTstop-profiling/03_PeakDifferencesFinal/data/constructInfo2.rds")


pqsDefaultRes <- readRDS("/Users/mariokeller/projects/HNRNPH_project/RTstop-profiling/rds/pqsDefaultRes.rds")
pqsDefaultRes <- pqsDefaultRes[!seqnames(pqsDefaultRes) %in% 12001:12010]
pqsScoreThreshold25Res <- readRDS("/Users/mariokeller/projects/HNRNPH_project/RTstop-profiling/rds/pqsScoreThreshold25Res.rds")
pqsScoreThreshold25Res <- pqsScoreThreshold25Res[!seqnames(pqsScoreThreshold25Res) %in% 12001:12010]
qgrsDefaultRes <- readRDS("/Users/mariokeller/projects/HNRNPH_project/RTstop-profiling/rds/qgrsDefaultRes.rds")
qgrsDefaultRes <- qgrsDefaultRes[!seqnames(qgrsDefaultRes) %in% 12001:12010]
```


# Workflow

I first determined all constructs with no predicted G4 in the three approaches (pqsfinder default, pqsfinder score threshold of 25 (less strigent) and QGRS Mapper default). Afterwards, I first determined all significantly upregulated peaks (padj $\le$ 0.05 and log2FC > 0) that are located on constructs where all significant peaks are upregulated (dirStatus == All up) and subsequently filtered these peaks to those located on a construct without G4 prediction.

```{r constructsWithNoPredictionButMeasuredG4}
withPrediction <- c(seqnames(pqsDefaultRes) %>% as.character, seqnames(pqsScoreThreshold25Res) %>% as.character, seqnames(qgrsDefaultRes) %>% as.character) %>% unique
withoutPrediction <- setdiff(as.character(sprintf("%05d", 1:12000)), withPrediction)

#get upregulated peaks
upregPeaks <- peaksChange[peaksChange$log2FoldChange > 0 & peaksChange$padj<=0.05 & peaksChange$dirStatus=="All up"]
#subset to upregulated peaks from constructs with no prediction
upregPeaksWithoutPrediction <- upregPeaks[seqnames(upregPeaks) %in% withoutPrediction]

kable(
  data.frame(cat=c("Constructs with predicted G4", "Constructs without predicted G4", "Upregulated peaks", "Upregulated peaks and no prediction"),
             n=c(length(withPrediction), length(withoutPrediction), length(upregPeaks), length(upregPeaksWithoutPrediction)))
)
```

# Output 

Results are stored in upregulatedPeaksWithoutPrediction.xlsx 

```{r output}

upregPeaksWithoutPrediction$regionID <- constructInfo2$Region.ID[match(upregPeaksWithoutPrediction$constructID, constructInfo2$Construct.ID)]
upregPeaksWithoutPrediction$geneName <- constructInfo2$Gene.Name[match(upregPeaksWithoutPrediction$constructID, constructInfo2$Construct.ID)]

output <- upregPeaksWithoutPrediction %>% as.data.frame() %>% tibble::rownames_to_column(., var="peakID") %>% mutate(., seqnames = as.character(seqnames)) %>% 
  dplyr::select(., constructID, regionID, geneName, peakID, start, end, baseMean, log2FoldChange, padj, expressionStatus, dirStatus, highConf) %>% arrange(., desc(log2FoldChange))

xlsx::write.xlsx(output, file="/Users/mariokeller/projects/HNRNPH_project/RTstop-profiling/XLSX/upregulatedPeaksWithoutPrediction.xlsx", row.names=FALSE)

```

\pagebreak 

# Session Information

```{r}
sessionInfo()
```
